---
title: "Analysis of bikeshare data"

author: "Cameron Mitchell"
date: "11/6/2020"
output: html_notebook
---

The goal of this analysis is to provide some prelimiary analysis on the bike share data for August 2020. I will be focusing on finding the deficits and gains for each of the bikestations

## Prepare workspace:

#### Load packages

We will work primarily within the context of the tidyverse set of packages, with a few additional packages supporting exploratory analysis.  I have included the `lubridate` package, as we will want to do some work with dates.

```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(skimr)
library(summarytools)
library(lubridate)
```


## Data


The orginal source of the data was the csv file 

https://s3.amazonaws.com/capitalbikeshare-data/202008-capitalbikeshare-tripdata.zip

so I will use that original source data so that my research is more *reproducible*.


### Read the data

I read the original .csv file and add in the variables  `duration`, `hour_of_day`, and `day_of_week`.

```{r}
dfa= read_csv("../Data/202008-capitalbikeshare-tripdata.csv") %>% 
  clean_names() %>%
  mutate(duration=as.numeric((ended_at-started_at)/60),
         hour_of_day=hour(started_at),
         day_of_week=wday(started_at,label = T))
  
```

We summarize that dataframe to identify data types, missing data, et cetera.

```{r}
skim(dfa)
```

#### A histogram - like we did in class

As we discussed in class, we might be interested in looking at histogram of durations, broken out for each day of the week.  Here I filter to positive duration shorter than 100 minutes.

```{r}
dfa %>% filter(duration<100,duration>0) %>% 
  ggplot(aes(x=duration)) + geom_histogram(bins = 300)+facet_wrap(~day_of_week)
```


### Cleaning

Limit to durations that are positive.

```{r}
dfb=dfa %>% filter(duration>0)
```

### This section will be used to determine the number of bikes gained or lost at each bike station.

Create a dataframe that for each bike station it contains the number of starts and ends and then compute the net gain.

```{r}
# number of starts at each station
dfstart = dfb %>% group_by(station = start_station_id) %>%
  summarise(start_num = n())

# number of ends at each station
dfend = dfb %>% group_by(station = end_station_id) %>%
  summarise(end_num = n())

# joining the two dataframes with all the valid data and computing the net gain
dfc = merge(dfstart, dfend) %>%
  filter(!is.na(station)) %>%
  mutate(net_gain = start_num - end_num)
```
The code below is just another way of getting a similar dataframe that I might want.
```{r}
dfriders = dfb %>%
  pivot_longer(started_at:ended_at, names_to="type",values_to="t") %>% arrange(t) %>%
  mutate(increment=case_when(
   type=="start_date"~1,
   type=="end_date" ~ -1)) %>%
  mutate(riders=cumsum(increment)) %>%
  group_by(station = start_station_id) %>%
  summarise(start_num = n())
```

### Visualization

Using ggplot2 I will try to show some of the results

```{r}
dfc %>% ggplot(aes(x = net_gain)) + geom_histogram()
```

This is the only graph that I could think of creating with dfc and it shows that most of the stations have a net_gain around zero with a few drastic outliers