---
title: "Analysis of bikeshare data"

author: "Cameron Mitchell"
date: "11/6/2020"
output: html_notebook
---

The goal of this analysis is to provide some prelimiary analysis on the bike share data for August 2020. I will be focusing on finding the deficits and gains for each of the bikestations

## Prepare workspace:

#### Load packages

We will work primarily within the context of the tidyverse set of packages, with a few additional packages supporting exploratory analysis.  I have included the `lubridate` package, as we will want to do some work with dates.

```{r}
library(tidyverse)
library(janitor)
library(readxl)
library(skimr)
library(summarytools)
library(lubridate)
```


## Data

I am **not** using the file that (a) started on moodle for which we,  (b) computed a couple of extra variables (for duration, hour of day, and day of week), and (c) saved as .xlsx file.  

The orginal source of the data was the csv file 

https://s3.amazonaws.com/capitalbikeshare-data/202008-capitalbikeshare-tripdata.zip

so I will use that original source data so that my research is more *reproducible*.

>Note 1: My variable names might be different than yours.  Be careful to think through what I am doing and use it as a template.




### Read the data

I read the original .csv file and add in the variables (mutate) that we did manually when we played with the creation of the data table.   I will call these new variables: `duration_min`, `hour_of_day`.  Also I will add `day of week`, but it won't match the format from excel.

```{r}
dfa= read_csv("202008-capitalbikeshare-tripdata.csv") %>% 
  clean_names() %>%
  mutate(duration=as.numeric((ended_at-started_at)/60),
         hour_of_day=hour(started_at),
         day_of_week=wday(started_at,label = T))
  
```


We summarize that dataframe to identify data types, missing data, et cetera.

```{r}
skim(dfa)
```

#### A histogram - like we did in class

As we discussed in class, we might be interested in looking at histogram of durations, broken out for each day of the week.  Here I filter to positive duration shorter than 100 minutes.

```{r}
dfa %>% filter(duration<100,duration>0) %>% 
  ggplot(aes(x=duration)) + geom_histogram(bins = 300)+facet_wrap(~day_of_week)
```


### Cleaning

Before we do further analysis, we recognize that if the duration is negative, some piece of the time data is corrupt.  Although it sill reduce the size of our dataset, it is very much large enough to permit continued analysis even after removing those rows.

Limit to durations that are positive.

```{r}
dfb=dfa %>% filter(duration>0)
```

### This section will be used to determine the number of bikes gained or lost at each bike station.

Create a dataframe that for each bike station it contains the number of starts and ends and then compute the net gain.

```{r}
# number of starts at each station
dfstart = dfb %>% group_by(station = start_station_id) %>%
  summarise(start_num = n())

# number of ends at each station
dfend = dfb %>% group_by(station = end_station_id) %>%
  summarise(end_num = n())

# joining the two dataframes with all the valid data and computing the net gain
dfc = merge(dfstart, dfend) %>%
  filter(!is.na(station)) %>%
  mutate(net_gain = start_num - end_num)
```

### Visualization

Using ggplot2 I will try to show some of the results

```{r}
dfc %>% ggplot(aes(x = net_gain)) + geom_histogram()
```

This is the only graph that I could think of creating with dfc and it shows that most of the stations have a net_gain around zero with a few drastic outliers


